Estimating the likelihood of student replication success after an initial failure to replicate

### 2024-03-08 manuscript/manuscript.Rmd
welp, I wanted to put citations in the tables, but that only works in format markdown and the headers only work with format latex ugh
a rough editing pass over everything

---
title: "Estimating the likelihood of student replication success after an initial failure to replicate"
author:
  - name: Veronica Boyce
    affiliation: Stanford
    footnote:
      - corresp
  - name: TODO
    affiliation: Stanford
  - name: Michael C. Frank
    affiliation: Stanford
address:
  - code: Stanford
    address: Stanford University
footnote:
  - code: corresp
    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251rescue.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: apa6.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt

@@

As a pragmatic matter, when a replication fails, scientists have to decide: do they make a second attempt at replication, or move on?

We re-replicated 17 experiments that had previously had a failed student replication, reported in Boyce et al. (2023). In 5/17 of these rescue projects (29%), the results mostly or fully replicated. We discuss the relative effect sizes and levels of statistical consistency between original effect sizes and replication effect sizes.

While a few studies with failed replications were rescued with larger sample sizes and experimental tweaks, our overall results indicate that most failed replications stayed failed. If the goal is to do cumulative science, one may wish to stop after one failed replication.

@@

imagine a scenario when this happens (motivating example). Imagine the scenario where you are a new graduate student, and you run a replication of a study in the literature that you are interested in building upon in your research. The replication fails. TODO elaborate -- some hints of the effect yo hoped for or maybe the effect is close to null. (or maybe the pattern is off even if one part is suggestive)  Should you try again, running another replication, perhaps with a larger sample or other tweaks to possibly fix the replication? Or should you give up on replicating that specific study, and find a different study to replicate and build upon?

we focus on scientists’ decisions… (spelling out why they might be replicating and hence why they might want to re-replicate) 
TODO maybe instead of starting with p(true), start with “We focus on scientists’ decisions about what studies to run and how to expend their limited resources in the face of an uncertain literature.” THEN you can come back to “this perspective differs from the standard approach to replication, which has been concerned with the proportion of true effects in the literature. We take a different angle on replications, and instead focus scientists decisions on what studies to run. Individual studies (and replications) are rarely definitive when it comes to scientific evidence, but they often influence what studies a scientist conducts.  If a study works, scientists are likely to continue to purse that result, but if a study doesn't work, scientists have to choose whether to try again, or switch to a different study.

this situation comes up a lot because replication is an important issue. (review meta-science literature on replication - you have to do this)

replications can fail for many reasons (move up this material) The idea of replication failure implies a certain time direction: first a "successful" study and (later) a "failed" replication. (CITE TIME INVARIANCE) We can also consider this as two very-related studies that got different results, and thus the question is what difference between these two studies could account for the differing results.

[not sure if it would be better to structure this differently]

A replication can have different results from the original study for a number of different potential reasons (see also @ebersole2020). The original result may have been a fluke, through chance or p-hacking. The original result might be of much less generalizability than is assumed, and very sensitive to the exact conditions and time it was run in ("hidden moderators"). The original could have an inflated effect size, due to some combination of heterogeneity, p-hacking, luck, and publication bias.

The replication could have been underpowered due to any of inflated original effect size, attrition in sample, or the original also being underpowered. A manipulation might have failed in the replication (this is a proximate cause that also raises the question of why the manipulation check failed). There might have been a difference in materials or instructions between the two, that turns out to matter (more likely if materials are not available). The experiment might not port well to online due to some details. A change in sample (ie. to online) may decrease data quality without sufficient attention checks. The replication might have made changes (for convenience or a new sample) that turn out to matter.

[ maybe try to make the point here that if there are changes in time/place/population, you want a good *translation* of the experiment not an identical copy]

While one can theorize after the fact about potential causes, in most circumstances it is hard to diagnose for certain what made for the difference in the results. An exception might be if a specific error in one study or the other is found, such that the results were not what was thought (analytic error) or the data was not what was thought (experimental error).

but re-replication has not been studied as much (except MB5 - use this paragraph to describe) How definitive is a replication failure? That is, if a study fails to replicate, what are the chances that an additional replication succeeds. After CITE OSF found an overall replication rate of WHATEVER, concerns were raised by @gilbert that method differences were the problem. In 11 of the OSF studies, there had been concerns about protocol fidelity raised by original authors prior to data-collection. 10 out of these 11 failed to obtain significant results in the replication. In a follow-up, @ebersole2020 re-replicated 10 of these 11 in large samples, using both the RP:P protocol and a new protocol revised under advise of the original authors or other experts. None of the RP:P protocols found significant results, 2 of the 10 revised did (but not the 1 that was significant in the original RP:P sample). The replication effect size was much smaller CITE NUMBER. We could frame @ebersole2020 as an attempt to rescue the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks.

re-replication could be a strategy to implement best practices that have been hypothesized to be positive (might need a paragraph here on sample size for replications, can use the references in Experimentology sampling chapter, there is a box on this topic)

<!--There has been much discussion in the literature about how we should interpret replication results in terms of the existence of a true effect or effect size. TODO CITATIONS-->

<!--Well, it depends on what you're trying to measure. From the perspective of learning from the field as a whole things are complicated and may depend also on a priori information from other related studies, the beleivability of the results, the relative rigor and sample sizes of the original and replication (TODO perhaps cite the time invariant). -->

What would these potential causes of replication failure suggest would happen in another replication? For our purposes, we're considering another replication with roughly the same environment, knowledge, and resources of the first replication, but with the benefit of hindsight on the first replication.

Heterogeneity and lack-of-generalizability reasons mostly suggest another failed replication, as the second replication is likely to be closer to the first replication than to the original in these random factors. CITE THAT PAPER about types of heterogeneity and which way they would fall

If the original result was p-hacked or a luck false positive, then another replication is also likely to fail.

Manipulation check failure may not be fixable, but it might at least be diagnosable.

Underpoweredness and inflated effect size may be recoverable depending on the extent -- if the effect size was inflated and the first replication powered for the effect, a better powered (but not huge) replication may be able to recover a smaller, but still convincing effect.

Changes between original and replication may or may not be fixable -- if the original materials are still unavailable, a second replication is no better off. If a second replication can get closer to the original, either by using closer materials, or better adapting the study to a new environment, the rescue might succeed at better approximating the results of the original.

Sensitivity to environment or not working well online are unlikely to be fixable if the replications remain constrained to using online methods. Similarly, if the result is specific to a special population, then a new replication is also likely to fail, unless it has gained access to this population.

<!--These are pragmatic judgments related to how a lab might treat a failed replication as a sign that they cannot do cumulative science on a particular paradigm. In some cases, especially related to changes-in-population or changes-to-online, replications, and thus cumulative work on the paradigm, might be possible for other groups who have different resources.

The question is, having a failed replication, and not knowing which reasons might have caused the different results -- should a scientist try again? -->

Overall, it seems that there are both potential causes of replication failure that may be resolvable even with a similar level of resources and potential causes that are not resolvable, either at all, or given resource constraints.

the current study

[Probably needs a better transition in] Here we report the results of `r num_proj` re-replication "rescue" projects which each re-replicated a study which had a failed or only partially-successful replication in @boyce2023. This is a non-random sample, and given the small number of projects, we provide descriptive statistics and discuss qualitatively some potential sources of differential re-replication outcomes.

@@

A spreadsheet of projects, individual project write-ups (both replications and rescues), links to individual project data and analyses for rescue projects, and the analytic code for this paper are all available at OSF LINK GOES HERE.

@@

Students were required to think of reasons the original replication might not have worked, and address them if they could. A list of possible reasons and solutions TODO LINK was given to students. In general, we encouraged students to add manipulation checks as appropriate, and better adapt materials to online studies. For instance, TODO rescue switched from using the CRT which is overused in online samples for the newer TODO. TODO rescue discovered that the replication had accidentally used the drawn version of the stimuli rather than the photographic stimuli used in the original and reverted to the photographic stimuli. TODO @tarampi original had participants indicate their answers (left or right) on a piece of paper in a timed navigation task, the replication had them indicate by clicking a drop down, and rescue went with press keyboard keys to indicate. TODO others to note?  Once students experimental designs and analytic plans were approved by TAs (VB and BP), students pre-registered and ran their samples.

With one exception, samples were collected on Prolific (the rescue of @yeshurun2003 ran in-person on the Stanford student subject pool). We tried to power studies adequately (with a target of 2.5x original following @simonsohn2015), but due to cost constraints, not all studies were powered at this level. The rescues had on average `r round(m,2)` times the original sample post-exclusion (median: `r round(mid,2)`, IQR: `r round(low,2)` - `r round(high,2)`, minimum: `r round(lowest, 2)`, maximum: `r round(highest,2)`, see TABLE TODO for all sample sizes). 
Across the `r num_proj-1` Prolific studies, we spent $`r pluck(cost,1,1) |> round()`, for an average of $`r (pluck(cost,1,1)/(num_proj-1)) |> round()` per project.

@@

Another successful replication was that of @ngo2019. Here, the original study found a large effect, and so the first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people, and then failed to find the effect. The rescue, powered using 2.5x the original sample (as recommended by @simonsohn2015a), recovered a clear effect (albeit a much smaller one). There are reasons to think that some effect sizes in the literature may be inflated TODO CITATIONS, and separately potential reasons that slight changes to experiments, or switches to online, could result in noisier samples (and thus smaller effect sizes). Thus, replications with smaller samples than the original (even if powered to the original effect size), may not be that diagnostic, and could potentially benefit from a re-replication.

@@

@@

All of the correlations are presented in Figure TODO. As the number of rescues is small, and many of these predictors are correlated, we caution against over-interpretation. The strongest correlates of rescue success were open materials, a small sample size on the first replication, a small sample size on the first replication relative to the original sample size, and a large rescue sample size relative to the first replication. None of these effects meet the conventional significance threshold.

Small replication samples relative to original and rescue could be due both to a) aiming for a small replication sample due to aiming for power for a reported large effect size or b) difficulties with recruitment or high exclusion rates leading to a smaller than intended sample. Since relative sizes of the studies may play a role in replication success and how probitive replications are, we show the sample sizes in TABLE TODO.

An additional factor that influences the interpretation of a replication is how close the replication's methods were to the original. In the rescue projects, we aimed to have methods be as close as was feasible or appropriate. (As discussed above, changes in time, population, or setting may necessitate adaptation of a study to maximize fidelity.) However, rescue projects varied in how close the re-replications actually were, often due to limitations in the availability of original stimuli and original instructions, in addition to the use of primarily online subject pools. TABLE TODO shows the closeness of each re-replication according to the classification scheme from @lebel.

```{r}
parsed_d |> select(name_pretty, type, N, closeness, replication_score) |> 
  filter(type!="additional") |> 
  pivot_wider(names_from="type", values_from=c(N, closeness, replication_score)) |> 
  select(paper=name_pretty, replication_score_rescue, N_original, N_rep1, N_rescue,
         "Replication \\\\  closeness"=closeness_rep1, closeness_rescue) |> arrange(replication_score_rescue|> desc()) |> knitr::kable(align="lllll", format="latex", booktabs=T, linesep="", caption="Comparison of sample size for original, replication, and rescue samples and measures of closeness for replication and rescue samples.", col.names=c("", "", "Original", "Replication", "Rescue", "Replication", "Rescue")) |> add_header_above(c("Paper"=1, "Score"=1, "N"=3, "closeness"=2))

@@

```{r, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. "}

@@

```{r, include=F, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. "}

@@

We standardized all effect sizes into standardized mean difference (SMD) units. One potential issue with comparisons using SMD is that noisier measures will have smaller standardized effect sizes even if the effect on the original scale is the same. In general, the replication and rescue effect sizes were smaller than the original effect sizes, and in a couple cases the effects were in the opposite direction (FIGURE TODO).

@@

```{r}
names <- parsed_d |> select(target_lastauthor_year, name_pretty) |> unique()

@@

Here, we measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that two effects come from the same distribution. As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found). Because we have few replication for any study, we impute the heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson2020 in prior multi-site replications in psychology.

The median value of $p$-original between an original study and its replications was `r main_summ$median |> round(2)` [IQR: `r main_summ$low |> round(2)` - `r main_summ$high |> round(2)`]. `r round(main_summ$less_than*100)`% of the $p$-original values were less than .05, indicating by conventional thresholds a rejection of the null hypothesis that the original and the replications came from the same distribution for the given imputed level of heterogeneity. Individual $p$-original values for each study are shown in TABLE TODO.

@@

# Discussion
We presented the results of `r num_proj` new replications that attempted to "rescue" previous failed replications reported in @boyce2023 by identifying possible causes of non-replication and ameliorating them. `r nrow(success)` of these replications (`r round(nrow(success)/num_proj*100)`%) mostly or fully replicated.

We don't have qualitative or quantitative explanations for why some replicated and some didn't. In a couple cases, increasing sample size and fixing internal validity issues in the replication seems to have led to a successful rescue (although we can't establish causality even in these cases). However, there were other studies that had small replication samples, or implementational deviations in the replication, and rescues were still unsuccessful. We can't predict what replication failures are likely to resolve given another try (or a more thoughtful try), beyond that suggestion that glaring problems and low samples may sometimes (but not always) be resolvable.

Another pattern we observed was that the effect sizes of even the successful replications tended to be substantially smaller than the original effect.

[SAY something about statistical consistency]

## Limitations
The reported rescue projects are a small sample of replications, and the effects explored are non-random, as they have been doubly selected by student interest. However, this non-random may be a useful selection bias as it is correlated with  how students choose what to work on.

The authors of the rescue projects put substantial effort into trying to set up rescues that had a good chance of success, but projects were constrained by budget limitations, a short timeline, and primarily running online studies. These limitations are representative of the sort of resource limitations often faced by early-career researchers. That said, it is possible that different results might be obtained in better-resourced settings. Thus, we do not make statements about whether these studies are "true" or "false-positives", merely that they do not support cumulative research by early-career researchers under these conditions.

We opened this paper with a question about what an early-career researcher should do given a failed replication. Should one try again or move on? From our sample of testing the "try again" approach, it seems that the odds of a re-replication working are low (consistent with @ebersole2020), so especially if there isn't a super clear failure mode to point to, another try probably won't fix it.

>>>>

---
title: "Estimating the likelihood of student replication success after an initial failure to replicate"
author:
  - name: Veronica Boyce
    affiliation: Stanford
    footnote:
      - corresp
  - name: TODO x a lot
    affiliation: Stanford
  - name: Michael C. Frank
    affiliation: Stanford
address:
  - code: Stanford
    address: Stanford University
footnote:
  - code: corresp
    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251rescue.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: apa7.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt

@@

As a pragmatic matter, when a replication fails, scientists have to decide: do they make a second attempt at replication, or move on? 
Given the relatively low rates of replication success reported in large scale replication projects, this choice point is one facing many scientists in psychology who attempt to replicate a study before building on it. Here, we report on 17 re-replications of experiments with failed replicated reported in  Boyce et al. (2023). In 5/17 of these rescue projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes. While a few studies with failed replications were rescued with larger sample sizes and experimental tweaks, our overall results indicate that most failed replications stayed failed. If the goal is to do cumulative science, one may wish to stop after one failed replication.

@@

imagine a scenario when this happens (motivating example). Imagine the scenario where you are a new graduate student, and you run a replication of a study in the literature that you are interested in building upon in your research. The replication fails: perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a p-value of .3. Or perhaps the interaction is numerically in the wrong direction, and also the main effects look different. Or perhaps the main effect of interest is null, but you found a significant effect on the interaction that wasn't hypothesized. Whatever the details of the replication failure, you are left with a question: Should you *try again* and run a re-replication, or should you *give up* and pick a different study to replicate and build on?

We focus on scientists’ decisions about what studies to run and how to expend their limited resources in the face of an uncertain literature. Scientists may replicate studies for a variety of reasons, including calibrating their equipment, confirming a result before building on it, and engaging in meta-scientific practices [other reasons to highlight? is there any citation for this?].
We focus on this middle motivation, where a scientist wants to build on a result, and thus wants it to succeed, but does not have strong reasons to believe or disbelieve the finding: it isn't a robust, well-tested effect, nor is it something they have particular suspicions about. However, in an uncertain literature, we might not want to commit resources to building on an effect without first checking that we, with our resources and participant sample, can get the core effect we wish to build on. Thus, scientists may want to re-replicate failed replications if the re-replication is likely to work and is indicative that future experiments in this tradition are likely to be informative.

This scientist-decision focused angle is different from the standard approach to replication which has take then the meta-scientific approach which has been primarily concerned with estimating the proportion of true effects in the literature. Large-scale replication studies indicate that the literature is unreliable.

In psychology, large scale replication projects have tended to find replicability rates around half. Across 100 studies with positive results, RP:P replicated 36 - 47% depending on the metric for replication success [@openscienceconsortium].  With large multi-site samples, Many Labs 1 replicated 11/13  effects (84.6%), Many Labs 2 replicated 14/28 effects (50%), and Many Labs 3 replicated 3/10 effects (30%) [@klein2014; @klein2018; @ebersole2016]. @camerer2018 included 21 behavioral social science studies and replicated 12-14 of them (57%-67%) depending on the metric used. @boyce2023 reports an average replication score of 49% for 176 replications primary in psychology. Psychology is not the only discipline where not everything replicates. Large scale replication projects in other disciplines have found 39/97 (40%) for positive effects in cancer biology [@errington2021], 11/18 (61%) in economics [@camerer2016], and 31/40 (78%) for experimental philosophy [@cova2021].

TODO DISCUSSION OF THE LITERATURE
What these replication rates say about the truth of the literature is beyond our scope, but these rates do indicate that scientists are far from guaranteed at being able to replicate effects from the literature in their own labs. Thus, the question of what to do with a failed replication is a practical one that comes up frequently.

Replications can fail for many reasons. Replication failure implies a certain time direction: first a "successful" study, usually one that got a certain statistically significant result, and (later) a "failed" replication that did not get this statistically significant result. The replication failure framing implies a time direction (TODO THERE IS A CITATION FOR THIS IDEA RIGHT OF PRECEDENCE). Taking out the time component, we can instead analyse a replication failure as two very-related studies with differing results. Under this frame, the question becomes: what is different between these two studies that could account for the differing results?

[query: should I try to frame this as two that are symmetric , or should I stick to an original / rep framing? I'm torn]

There are many potential reasons for differing results (see also @ebersole2020). One option is that one of the studies had a statistically unlikely result (either through chance or p-hacking). The result could be sensitive to some of the exact conditions and time when the two studies were run ("hidden moderators") and lack generalizability. Reported effect sizes might be biased estimates due to some combination of heterogeneity, p-hacking, luck, underpowering, and publication bias. One or both studies might be underpowered, potentially due to believing the effect was larger, or unexpected attrition in the sample.

Studies might differ in how they instantiated the question of interest, and this might cause different results. For example if a manipulation check failed, that could be a proximate cause of difference, which also raises questions about why the manipulation check failed. Differences in materials or instructions might turn out to be crucial to the effect. A difference in experiment platforms (ex. online versus in-person) might mean that an implementation doesn't fit the platform, or the data quality controls and attention checks may not be appropriate to the sample population. Some trivial changes intended to adapt the experimental question to the setting might actually change the result. If there are differences in time, place, or subject population between the two studies, there may need to be differences in the materials, instructions, or procedure to provide a good *translation* -- too many changes could cause differing results, but so could not enough changes.

While one can theorize after the fact about potential causes of replication failure, in most circumstances it is hard to diagnose for certain what made for the difference in the results. An exception might be if a specific error in one study or the other is found, such that the results were not what was thought (analytic error) or the data was not what was thought (experimental error). 
Re-replication could serve as a way of triangulating on reasons for replication failure, as well as another chance for a scientist to get a working protocol that picks up the effect of interest. However, re-replication has not been studied much, so it is uncertain how definitive a replication failure is. That is, conditioned on a failed replication, how likely is a re-replication to succeed? Re-replication was tested in a sample of 10 psychology studies in @ebersole2020. After @openscienceconsortium published their results with a replication rate of TODO, @gilbert2016  raised concerns that methodological differences were the problem. In 11 of the OSF studies, concerns about protocol fidelity had been raised by original authors prior to data-collection (and thus not contingent on results). Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication [@openscienceconsortium]. In a follow-up, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts, thus testing whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications. The result: 0 of the RP:P protocols found significant results, 2 of the 10 revised protocols did (but not the 1 that had a significant result in the original RP:P sample). The replication effect size was much smaller CITE NUMBER [@ebersole2020]. We could frame @ebersole2020 as an attempt to rescue the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks. This suggests a successful re-replication rate of 2/9 (22%) for studies with non-significant first replications, under fairly favorable, high resource circumstances.

Returning to the question of potential causes of non-replication, what do reasons for replication failure say about the success of another replication? For our purposes, we're considering another replication with roughly the same environment, knowledge, and resources of the first replication, but with the benefit of hindsight on the first replication, as this is the situation scientists considering doing a re-replicaton for cumulative science are in.

Factors that suggest another failure include heterogeneity and lack-of-generalizability ("hidden moderator"), as the second replication is likely to be closer to the first replication than to the original in these random factors.  TODO CITE THAT PAPER about types of heterogeneity and which way they would fall An original result due to p-hacking or a lucky false positive is unlikely to re-replicate. Sensitivity to environment or not working well online are unlikely to be fixable if the replications remain constrained to using online methods. Similarly, if the result is specific to a special population, then a new replication is also likely to fail, unless it has gained access to this population.

Factors that may be fixable include sample size issues with the replication, assuming the re-replication can recruit a larger sample or reduce attrition. It may be possible to fix underpoweredness and inflated effect size, at least to an extent. If the effect size was inflated and the first replication powered for the effect, a better powered (but not huge) replication may be able to recover a smaller, but still convincing effect.  Failures due to differences between the materials may or may not be fixable. If the original materials are still unavailable, a second replication is no better off. If a second replication can get closer to the original, either by using closer materials, or better adapting the study to a new environment, the rescue might succeed at better approximating the results of the original. A failed manipulation check failure may not be fixable, but it might at least be diagnoseable.

Overall, there are some potential causes of replication failure that may be resolvable with a similar level of resources, and also some potential causes of replication failure that are not resolvable, either at all, or given resource constraints.

After a failed replication, we might speculate on what caused the replication result, and what that would mean for future work on the topic. As the @gilbert2016 and @ebersole2020 example showed, these speculations on causes and thus re-replication outcomes are not always correct.

While @ebersole2020 looked at re-replication success rate in a high resource setting with access to large, multi-site samples and expert input, most early career scientists do not have those kinds of resources to devote to re-replication attempts. On the other hand, early career researchers may also be prone to a different distribution of causes of initial replication failure.

Here, we investigate re-replication in the context of graduate student replications. Students in a graduate methods class on experimental methods were learning about best practices for experimental research. One option for their class project was to "rescue" an experiment that a student in a prior year had previously failed to replicate (reported in @boyce2023). Thus, students were likely to be looking at both the original paper and the failed replication through the lens of best practices to see what they could fix. [TODO Mike mentioned something about a sampling side box on experimentology that VB is having trouble finding]

In the present paper, we report the results of `r num_proj` re-replication "rescue" projects which each re-replicated a study which had a failed or only partially-successful replication in @boyce2023. This is a non-random sample, and given the small number of projects, we provide descriptive statistics and discuss qualitatively some potential sources of differential re-replication outcomes. [possibly drop last sentence?]

@@

A spreadsheet of projects, individual project write-ups (both replications and rescues), links to individual project data and analyses for rescue projects, and the analytic code for this paper are all available at TODO OSF LINK GOES HERE.

@@

Students were required to think of reasons the original replication might not have worked, and address them if they could. A list of possible reasons and solutions TODO LINK was given to students. In general, we encouraged students to add manipulation checks as appropriate, and better adapt materials to online studies. For instance, the rescue of @paxton2012 switched from using the CRT which is overused in online samples for the newer TODO. The rescue of @jara-ettinger2022 discovered that the replication had accidentally used the drawn version of the stimuli rather than the photographic stimuli used in the original and reverted to the photographic stimuli. @tarampi2016 original had participants indicate their answers (left or right) on a piece of paper in a timed navigation task, the replication had them indicate by clicking a drop down, and rescue went with press keyboard keys to indicate. TODO others to note?  Once students experimental designs and analytic plans were approved by TAs (VB and BP), students pre-registered and ran their samples.

With one exception, samples were collected on Prolific (the rescue of @yeshurun2003 ran in-person on the Stanford student subject pool). We tried to power studies adequately (with a target of 2.5x original following @simonsohn2015), but due to cost constraints, not all studies were powered at this level. The rescues had on average `r round(m,2)` times the original sample post-exclusion (median: `r round(mid,2)`, IQR: `r round(low,2)` - `r round(high,2)`, minimum: `r round(lowest, 2)`, maximum: `r round(highest,2)`, see Table \@ref(tab:tab-size) for all sample sizes). 
Across the `r num_proj-1` Prolific studies, we spent $`r pluck(cost,1,1) |> round()`, for an average of $`r (pluck(cost,1,1)/(num_proj-1)) |> round()` per project.

@@

Another successful replication was that of @ngo2019. Here, the original study found a large effect, and so the first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people, and then failed to find the effect. The rescue, powered using 2.5x the original sample (as recommended by @simonsohn2015), recovered a clear effect (albeit a much smaller one). There are reasons to think that some effect sizes in the literature may be inflated TODO CITATIONS, and separately potential reasons that slight changes to experiments, or switches to online, could result in noisier samples (and thus smaller effect sizes). Thus, replications with smaller samples than the original (even if powered to the original effect size), may not be that diagnostic, and could potentially benefit from a re-replication.

@@

```{r cor-table}

@@

All of the correlations are presented in Figure \@ref(tab:cor-table). As the number of rescues is small, and many of these predictors are correlated, we caution against over-interpretation. The strongest correlates of rescue success were open materials, a small sample size on the first replication, a small sample size on the first replication relative to the original sample size, and a large rescue sample size relative to the first replication. None of these effects meet the conventional significance threshold.

Small replication samples relative to original and rescue could be due both to a) aiming for a small replication sample due to aiming for power for a reported large effect size or b) difficulties with recruitment or high exclusion rates leading to a smaller than intended sample. Since relative sizes of the studies may play a role in replication success and how probitive replications are, we show the sample sizes in Table \@ref(tab:tab-size).

An additional factor that influences the interpretation of a replication is how close the replication's methods were to the original. In the rescue projects, we aimed to have methods be as close as was feasible or appropriate. (As discussed above, changes in time, population, or setting may necessitate adaptation of a study to maximize fidelity.) However, rescue projects varied in how close the re-replications actually were, often due to limitations in the availability of original stimuli and original instructions, in addition to the use of primarily online subject pools. Table \@ref(tab:tab-size) shows the closeness of each re-replication according to the classification scheme from @lebel2018.

```{r tab-size}
parsed_d |> select(name_pretty, type, N, closeness, replication_score) |> 
  filter(type!="additional") |> 
  pivot_wider(names_from="type", values_from=c(N, closeness, replication_score)) |> 
  select(paper=name_pretty, replication_score_rescue, N_original, N_rep1, N_rescue,
         "Replication \\\\  closeness"=closeness_rep1, closeness_rescue) |> arrange(replication_score_rescue|> desc()) |> knitr::kable(align="lllll", format="latex", booktabs=T, linesep="", caption="Comparison of sample size for original, replication, and rescue samples and measures of closeness for replication and rescue samples.", col.names=c("", "", "Original", "Replication", "Rescue", "Replication", "Rescue")) |> add_header_above(c("Paper"=1, "Score"=1, "N"=3, "closeness"=2))

@@

```{r smd, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. "}

@@

```{r , include=F, fig.cap="Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. "}

@@

We standardized all effect sizes into standardized mean difference (SMD) units. One potential issue with comparisons using SMD is that noisier measures will have smaller standardized effect sizes even if the effect on the original scale is the same. In general, the replication and rescue effect sizes were smaller than the original effect sizes, and in a couple cases the effects were in the opposite direction (Figure \@ref(fig:smd)).

@@

```{r tab-porig}
names <- parsed_d |> select(target_lastauthor_year, name_pretty) |> unique()

@@

Here, we measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that two effects come from the same distribution. As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found). Because we have few replication for any study, we impute the heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

[ query: should we bootstrap any of the following? ]

The median value of $p$-original between an original study and its replications was `r main_summ$median |> round(2)` [IQR: `r main_summ$low |> round(2)` - `r main_summ$high |> round(2)`]. `r round(main_summ$less_than*100)`% of the $p$-original values were less than .05, indicating by conventional thresholds a rejection of the null hypothesis that the original and the replications came from the same distribution for the given imputed level of heterogeneity. Individual $p$-original values for each study are shown in Table \@ref(tab:tab-porig).

@@

# Discussion
We presented the results of `r num_proj` new replications that attempted to "rescue" previous failed replications reported in @boyce2023 by identifying possible causes of non-replication and ameliorating them in the re-replication. `r nrow(success)` of these re-replications (`r round(nrow(success)/num_proj*100)`%) mostly or fully replicated.

We don't have qualitative or quantitative explanations for why the specific studies did or did not replicate. In a couple cases, increasing sample size and fixing internal validity issues in the replication seems to have led to a successful rescue (although we can't establish causality even in these cases). However, there were other cases where the original replications had issues with a small sample or deviations in the implementation, and the re-replication addressed these issues but still failed to replicate the original results. We can't predict what replication failures are likely to resolve given another, more thoughtful try, beyond that suggestion that glaring problems and low samples may sometimes (but not always) be resolvable.

Across all the re-replications, regardless of whether the pattern of effects replicated or not, the effect sizes were smaller than the original effect, often by a substantial amount. In a substantial minority of the replications, the replication effect sizes (taken as a whole) were statistically inconsistent with the original effect, even allowing for some heterogeneity between studies. These diminished and inconsistent effects suggest that even if a re-replication "works", it may be difficult to build upon as follow-up studies will need large samples to detect the small effects.

The reported rescue projects are a small sample of replications, and the effects explored are non-random, as they have been doubly selected for by student interest. However, this selection bias is likely to correlate with how graduate students choose what topics to work on and what studies to build on. Nonetheless, with a small sample all our estimates are highly uncertain.

The authors of the rescue projects put substantial effort into trying to set up rescues that had a good chance of success, but projects were constrained by budget limitations, a short timeline, and primarily running online studies. These limitations are representative of the sort of resource limitations often faced by early-career researchers. That said, it is possible that different results might be obtained in better-resourced settings, by scientists with more expertise and more time. Thus, we do not make statements about whether the original results are "true" or "false-positives", we merely claim they do not support cumulative research by early-career researchers under the constrained conditions we tested.

We opened this paper with a question about what an early-career psychology researcher should do given a failed replication: should they try again or move on? From our sample of testing the "try again" approach, it seems that the odds of a re-replication working are low (consistent with @ebersole2020), so especially if there isn't a super clear failure mode to point to, another try probably won't rescue the original results.
